---
title: "Subsetting"
author: "Abby Lewis"
date: "2025-06-02"
output: html_document
---

## Introduction

This worksheet is designed to help you practice a few different types of subsetting, which is often important when doing data analysis in R. Answers to all questions are at the bottom of the worksheet.

To start: run this chunk to load the data we will be working with

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Load data
data <- iris #iris is a dataset of flower measurements that is built into R
#Isolate one column to work with in the exercises below
mystery <- iris$Petal.Width
```

### Vectors

Q1) What is the 97th element in the vector `mystery`, which we created in the setup code above?

```{r}
#Add your answer here
```

Q2) Create a new vector `featured_numbers` that contains the 64th, 73rd, and 101st elements of `mystery`

```{r}
#modify this code:
featured_numbers <- mystery
featured_numbers
```

Q3) What happens if you use TRUE and FALSE to subset a vector? Predict the outcome of the code below before running it

```{r}
new <- c("Welcome", "to", "SERC")
new[c(TRUE, FALSE, FALSE)]
```

Q4) Create a new vector `large` that only contains the elements of `mystery` that are greater than 1.4

```{r}
#modify this code:
large <- mystery
large
```

Q5) How does your answer above work? Try explaining in words what is happening when you run that code

[answer here]

Q6) What happens if you use negative numbers to subset? Predict the outcome of the code below before running it

```{r}
new <- c(1, 2, 3)
new[-1]
```

Q7) Create a new vector `shortened` that contains all elements of mystery except for the first two. Use the function `length()` to confirm this is two elements shorter.

```{r}
#modify this code:
shortened <- mystery

length(mystery)
length(shortened)
```

### Data frames

Subsetting dataframes is similar to subsetting vectors, except that they have two dimensions

Q7) Isolate the third column of the iris dataset using its position (i.e., not the column name)

```{r}
#Answer
```

Q8) Isolate the fourth row of the third column of the iris dataset. You should get the value 1.5.

```{r}
#Answer
```

Q9) Now use the column name to isolate the column "Species" within the iris dataset. What is the 7th element (row) of this column?

```{r}
iris$Species
```

Q10) What if you realized that value was actually incorrect? Change this value (Species in the seventh row) to "virginica"

```{r}
#Your answer here

#Confirm that this worked
iris$Species[7]
```

## Nice work!

You made it through the worksheet! If you are interested in learning more about the nitty-gritty rules of subsetting in R, the Advanced R textbook from Hadley Wickham is a great resource

<https://adv-r.hadley.nz/subsetting.html>

# Answers

```{r}
#Q1
mystery[97]

#Q2
featured_numbers <- mystery[c(64, 73, 101)]

#Q4
large <- mystery[mystery > 1.4]

#Q5
#mystery > 1.5 produces a "logical vector", that is, a vector containing TRUE and FALSE. This vector is the same length as mystery. When you but this logical vector inside the brackets of mystery[] it tells R whether to keep (TRUE) or discard (FALSE) each element.

#Q7
iris[3]

#Q8
iris[4,3]

#Q9 
iris$Species
iris$Species[7]

#Q10
iris$Species[7] <- "virginica"
```
